---
title: "Lab 3"
author: "Zeyad Awad"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
library(statsr)
library(broom)
```

### Exercise 1

The dimensions of the dataset are 1458 rows x 123 columns. Each row represents an observation of a country in the HFI in a specific year.

```{r ex-1}
dim(hfi)
```

### Exercise 2

```{r ex-2}
hfi_2016 <- hfi %>%
  filter(year == "2016") %>%
  select(pf_expression_control, pf_score, year, countries, hf_score, pf_ss)
```

### Exercise 3

I would use a scatter plot to display the relationship between these two variables. The relationship looks linear. Hence, predicting the pf_score given pf_expression_control is possible if we use a linear model.

```{r ex-3}
ggplot(data=hfi_2016, aes(x = pf_expression_control, y = pf_score))+
  geom_point(alpha=0.7)

```

### Exercise 4

As depicted by the scatter plot in exercise 3 and the table below, there is a positive, strong correlation between these two variables. Additionally, there appears to be some outliers in the bottom left area of the graph.

```{r ex-4}
hfi_2016 %>%
  summarise(correlation = cor(pf_expression_control, pf_score))
```

### Exercise 5

The smallest sum of squares was 102.7. It is relatively close to my neighbours.

```{r ex-5}
plot_ss(x = pf_expression_control, y = pf_score, data = hfi_2016, showSquares = TRUE)
```

### Exercise 6
hf_score = 5.053396 + 0.368432 * pf_expression_control.
The slope tells us that for every unit increase in pf_expression_control, we expect, on average, a 0.368432 increase in a country's total human freedom score. In a nutshell, the less the political pressure on the media content, the higher the human freedom score.
```{r ex-6}
m1 <- lm(hf_score ~ pf_expression_control, data = hfi_2016)
tidy(m1)
glance(m1)
```
### Exercise 7
From the regression line, one could depict that for a country with a pf_expression_control of 3, their personal freedom score is predicted to be 5.906 (see table 3). However, the actual observation for this pf_expression_control value has a pf_score of 5.465632 (see table 2). Hence, this is an overestimate. The residual is -0.440368.  

```{r ex-7}
ggplot(data = hfi_2016, aes(x = pf_expression_control, y = pf_score)) +
  geom_point(alpha=0.7) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, color = "darkred")

hfi_2016 %>%
  filter(pf_expression_control == 3) %>%
  select(pf_expression_control, pf_score)

4.28+0.542*3
```
### Exercise 8
The appears to be no pattern in the residuals plot; they are scattered. This is an indicator that the two variables have a linear relationship.

``` {r ex-8}
m1_aug <- augment(m1)
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point(alpha=0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals")
```

### Exercise 9
Based on the histogram, the residuals are roughly normal-distributed - there is a little left-skewness that makes it not "perfectly" normal-distributed. Hence, the nearly-normal residual condition is not violated.

``` {r ex-9}
ggplot(data = m1_aug, aes(x = .resid)) +
  geom_histogram(binwidth = 0.25) +
  xlab("Residuals")
```

### Exercise 10
Based on the residuals vs fitted-values plot, there appears to be roughly constant variability, ie., no "major" fanning in or out. Hence, the constant variability condition is not violated.
``` {r ex-10}
m1_aug <- augment(m1)
ggplot(data = m1_aug, aes(x = .fitted, y = .resid)) +
  geom_point(alpha=0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  xlab("Fitted values") +
  ylab("Residuals") 

```

...
