---
title: "Lab 2"
author: "Zeyad Awad"
date: "`r Sys.Date()`"
output: openintro::lab_report
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(openintro)
```

### Exercise 1
As the binwidth gets smaller, the number of bins gets larger, so more features are revealed. For instance, a binwidth of 150 does not reveal the variation in counts as compared to a binwidth of 15. The default plot uses 30 bins.
```{r ex-1}
nycflights %>%
  ggplot(aes(x = dep_delay)) +
  geom_histogram()
nycflights %>%
  ggplot(aes(x = dep_delay)) +
  geom_histogram(binwidth = 15)
nycflights %>%
  ggplot(aes(x = dep_delay)) +
  geom_histogram(binwidth = 150)
```

### Exercise 2
There are 68 flights which are headed to SFO in February.
``` {r ex-2}
sfo_feb_flights <- nycflights %>%
  filter(dest == "SFO", month == 2)
sfo_feb_flights %>%
  summarise(counts = n())
```

### Exercise 3
The distribution is right skewed; hence we use the median and IQR because they are robust statistics. The median is -11 and the IQR is 23.2 as shown in the table.
``` {r ex-3}
sfo_feb_flights %>%
  ggplot(aes(x = arr_delay)) + 
  geom_histogram(binwidth = 25)

sfo_feb_flights %>%
  summarise(median_ad = median(arr_delay),
            iqr_ad = IQR(arr_delay))
```
### Exercise 4
DL and UA have the most variable arrival delays since they have the largest Interquartile Range (IQR = 22).
``` {r ex-4}
sfo_feb_flights %>%
  group_by(carrier) %>%
  summarise(median_ad = median(arr_delay),
            iqr_ad = IQR(arr_delay)) %>%
  arrange(desc(iqr_ad))
```
### Exercise 5
Using the median will be better in this case as the distribution is right-skewed, and the median is a robust statistic. However, the mean is useful if the distribution is symmetrical and it will be more useful in calculations. In this case, the month with the lowest median departure delay is September, whereas the month with the lowest mean departure delay is October.
``` {r ex-5}
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay),
            median_dd = median(dep_delay)) %>%
  arrange((median_dd))
nycflights %>%
  group_by(month) %>%
  summarise(mean_dd = mean(dep_delay),
            median_dd = median(dep_delay)) %>%
  arrange((mean_dd))

nycflights %>%
  ggplot(aes(x=dep_delay)) +
  geom_histogram(bins = 30)
```

### Exercise 6
I will choose LGA because it has the highest on-time departure percentage (~72.8%) as seen in the table.

``` {r ex-6}
nycflights <- nycflights %>%
  mutate(dep_type = ifelse(dep_delay < 5, "on time", "delayed"))

nycflights %>%
  group_by(origin) %>%
  summarise(ontime_dep_rate = sum(dep_type == "on time") / n()) %>%
  arrange(desc(ontime_dep_rate))

```
``` {r ex-x}
ggplot(data = nycflights, aes(x = origin, fill = dep_type)) +
  geom_bar()
```

### Exercise 7
We know that 1 mile per minute = 60 miles per hour.
``` {r ex-7}
nycflights <- nycflights %>%
  mutate(avg_speed = (distance/air_time)*60)
```

### Exercise 8
There appears to be a positive correlation. As the distance increases, the avg_speed increases. The trend looks to be logarithmic. Moreover, there appears to be some outliers beyond a distance of 3000.

``` {r ex-8}
nycflights %>%
  ggplot(aes(x = distance, y = avg_speed)) +
  geom_point(size = 0.7, alpha = 0.7, color="darkblue")
```

### Exercise 9
We can see that despite some flights take off late, they still arrive on time. The cutoff point is roughly 65 minutes (see graph 2).

``` {r ex-9}
nycflights %>%
  filter(carrier=="AA"|
           carrier=="DL"|
           carrier=="UA") %>%
  ggplot(aes(x = dep_delay, y = arr_delay, color = carrier))+
  geom_point()

nycflights %>%
  filter(carrier=="AA"|
           carrier=="DL"|
           carrier=="UA") %>%
  ggplot(aes(x = dep_delay, y = arr_delay, color = carrier))+
  geom_point() +
  geom_vline(xintercept=65, linetype = "dashed", size=0.7)+
  geom_hline(yintercept = 0, linetype = "dashed", size = 0.7)

```
...